\textbf{International Transport costs:\\New Findings from modeling additive costs}\\Online Appendix (Not for Publication)

### 2021-11-22 redaction/JEGeo/revision_JEGeo/referees/answers/Answer_referee1.tex

\date{September 2021}
 \maketitle
\bigskip

We would like to thank you for your insightful comments. They led us to introduce some
significant changes to the paper that we hope address your concerns. We first give you an overview
of the revision (Section \ref{sec:main_changes}) before answering in detail each of your comments (Section \ref{sec:detailed_answers}). Number of sections, pages and equations mentioned in the text refer to the revised version. When necessary, we refer to number of sections, pages and equations from the submitted version. In this case, they are written between brackets.

\section{Main changes \label{sec:main_changes}}

The structure of the paper has been modified after taking into account the referees' comments.
\begin{itemize}
\item In the submitted version, [Section 2] was devoted to the estimation of international transport costs; specifically, their break-down in two components, additive and ad-valorem. In [Section 3], we investigated the role of additive costs in the decomposition of transport costs time trends, between structural changes and composition effects. [Section 4] was devoted to the robustness analysis relative to the results from both previous sections.
\item In the revised version, we have strengthened the robustness checks regarding the estimation results of international transport costs (robustness checks to aggregation and to endogeneity have been added). Accordingly, the previously-called [Section 2] has been split in two Sections: Section 2, where we present the data and the estimation strategy; and Section 3, where we present the the results. Section 3 now includes the robustness checks as a final sub-section.
\item Both referees asked us to highlight the implications of additive costs (the ``Big Picture''). We agree that this was not enough emphasized in the submitted version, and we thank the referees for their suggestions on that issue. The revised version answers this concern in two points.
    \begin{enumerate}
    \item We maintain the analysis of the transport costs time trends (previously [Section 3], now Section 4); but we improved this section’s readibility by leaving the technical aspects in the Appendix (Section C). As such, we hope that the main message of this Section is easier to understand.
    \item Most importantly, we now emphasize the ``Big Picture'' welfare implications of additive costs on theoretical grounds. Through the lens of the \cites{melitz} model amended to integrate additive costs, we analyze the welfare gains that derive from the reduction in international transport costs estimated in Section 3. This allows us to quantify the extra welfare gains attributable to reduction of additive trade costs. This analysis forms Section 5.
        \end{enumerate}
\item To keep the paper (both the main text and with appendix included) within a reasonable number of pages, and stay fully transparent about our results, we have added numerous detailed tables to the Online Appendix.
\end{itemize}

@@

These three points indeed deserve careful consideration. We provide separate answers to each of them below.

%\subsubsection{On the justifications for our empirical approach}

\subsubsection{Concern (i): About the assumption of separability}
Our main empirical equation and its underlying assumptions regarding the separability of transport costs between their country- and product-level components draw on the one proposed by \citet{Irrazabal_2015} to estimate the share of additive costs in a firm-level context. It relies on a simple theoretical framework with minimal assumptions, and is compatible with most approaches within the so-called category of ``New Trade Theories''.

Further, we provide a robustness check for this separability assumption that $\tau_{ik} = \tau_i\tau_{k}$ and
 $t_{ik} = t_i +t_k$ in Section 3.3.1 of the paper (Table 2 and \textbf{Figure XXX}) \textbf{XXX CHECK AT THE END XXX}). We check the robustness of our results by re-running the estimation without the separability assumption. This comes at the cost of a substantial increase in the number of fixed in the regression. In 2019, for example, there are 14,016 country$\times$sector pairs. Without the separability assumption, one must estimate 28,032 fixed effects (number of country pairs $\times2$ for the two additive and multiplicative costs fixe effects) rather than ``only'' 836 ($(188+230)\times 2$). This is computationally impossible.
% use "/Users/guillaumedaudin/Documents/Recherche/2013 -- Trade Costs -- local/data/hummels_tra.dta", clear
% keep if year==2019
% keep if mode=="ves"
% gen sector=substr(sitc2,1,3)
% gen blouk = iso_o+sector
% codebook blouk
 For this reason, we have decided to run the robustness check on a reduced sample. For each year, we select the largest exporters that form together at least 80\% of annual trade and the largest traded sectors that form together at least 80\% of annual trade.
 We keep in the sample all trade observations from these exporters and in these sectors.
 This sample is smaller in terms of observations (2,125 for Air, 5,260 for Vessel on average over the period 1974-2019 (see Table 2), vs more than 30,000 on the complete sample, see Table 1). It yet remains quite large in terms of trade coverage (mean of 68\%).\footnote{In the restricted sample at the 3-digit level, we have 25 sectors from 14 countries for Air transport, vs 212 sectors from 190 countries in the baseline sample; and 54 sectors from 20 countries for Vessel, vs 667 sectors from 192 countries in the baseline sample). The vast majority of US imports comes from a selected range of countries, in a selected range of sectors.} On this reduced sample, we re-run the estimation with and without the separability assumption. The results are reported in Section 3.3.1 (Table 2 and \textbf{Figure XXX}). \textbf{XXX CHECK AT THE END XXX}) As we conclude at the end of this section, the trend pattern of the share of additive cost are very similar whether estimated under the separability assumption or not.
 % use "/Users/guillaumedaudin/Documents/Recherche/2013 -- Trade Costs -- local/results/comparaisons_various/stats_comp_baseline_non_séparé.dta", clear
% gen cover = cover_non_séparé/cover_baseline
% sum cover

@@

\begin{equation}
\ln f_{ikd} = \beta_{is(k)}\ln \tilde{p}_{ikd} + \text{Controls}_{ikd} +\epsilon_{ikd} \label{eq:estimation_ref1}
\end{equation}

\textit{where $d$ denotes the US district of entry and $k$ denotes an HS10 product} ($f_{ikd}$ being the transport costs). \textit{The
identification of $\beta_{ik}$, in this case, would rely on the across HS10 product and
district-of-entry variation in $f_{ik}$ and $p_{ik}$. Estimating the above equation would
obviously require that the authors do not aggregate up the raw Census data
across all districts and all 10-digit products pertaining to the same 5-digit category. [...] The first advantage of this so-called Hummels' approach is that the above regression can be estimated separately for various country-industry pairs, without imposing Assumptions (a) and (b) outlined above.}

\textbf{Our answer:} First of all, we thank the referee for suggesting to improve the comparison between our method and Hummels's one. We took this remark into account by adding a new sub-section in the revised version (Section 2.2, paragraph ``Estimation strategy'' \textbf{XXX Check XX}). As we now show in this Subsection, the elasticity of transport costs to unit prices $\beta_{is(k)}$ (in absolute value) also corresponds to the share of additive costs in total transport costs. The share of additive costs can hence be uncovered by regressing transport costs on unit fas prices, along the lines you suggest.

Now coming to the technical side, we thank the referee for his/her relevant suggestion of an alternative estimation method. This drove us to question our own empirical specification deeper. In the end, we decided to maintain our empirical strategy in the revised version, because its advantages seem to overweight its costs, and the alternative that the referee suggests does not go without a few important limitations. We develop each line of argument in our answer. Yet, because the referee's estimation method is also an interesting alternative, we have added a new section in the Online appendix devoted to it (Section D.3 \textbf{XXX CHECK XXX }).

@@

\paragraph{Estimating the share of additive costs: Highlighting some requirements} Our estimated equation relies on non-linear estimation methods, such as Non-Linear Least Squares. However, even with another formulation, such as the one suggested by the referee in Equation (\ref{eq:estimation_ref1}), we would still be constrained to resort to non-linear estimators. This is due to the necessity of imposing an \textit{ex-ante} restriction on parameters, i.e. $\tau \geq 1$ and $t \geq 0$, or $0 \leq  \beta \leq 1$. Should we relax these restrictions, standard linear, least squares estimates often deliver negative, meaningless estimates. In this respect, implementing the referee's method (see below) does not dispose of the requirement of resorting on non-linear estimates (and the computational, time-consuming burden it induces -- though the referee’s method is indeed less computational intensive). Imposing this parameters constraint was not made clear enough in the initial version, and we did our best to make this very important justification clearer in the revised version, as it is now exposed in footnote 12 page 8. \textbf{XXX Verifier à la fin? GD XXX }\smallskip

\paragraph{Exploring Referee's alternative functional form \label{subsec:functional_form}}

The estimation strategy suggested by the referee starts from Equation (\ref{eq:Hummels}) linking transport costs and unit prices as assumed in Hummels (2007), yielding Equation (\ref{eq:estimation_ref1}) as estimation equation. By year and transport mode, this implies running the estimation for each country of origin $i$ and each sector $s(k)$ = 3 digit sector, exploiting the variability between sub-sectors at the 10-digit level ($k$) and between ports of entry in the US ($d$). Despite its interest, the referee's method has some drawbacks, that in our view outweighs those of our method (in particular associated with \textit{Assumptions (a) $\tau_{ik} = \tau_i\tau_{k}$, and (b) $t_{ik} = t_i+ t_{k}$.}) Our answer can be articulated in two steps. First, we challenge the advantages that the referee's estimation method would bring in comparison with ours. Second, we emphasize the drawbacks induced by this method. \smallskip
%
%\paragraph{What we do}
%
%Our estimated equation is written as Equation (\ref{eq:equation0}) with $s$= 3 (or 4) digit and $k$=5 digit (based on Hummels' original dataset from 1974-2004). We thus exploit the variability of transport costs between countries and between 3-digit sectors (conditional on a year and a transport mode, air or vessel). To sum up our methodology:
%\begin{itemize}
%\item Estimating Equation \ref{eq:equation0} provides us with estimates of $\hat{t}_i$, $\hat{t}_{s(k)}$, $\hat{\tau}_i$, $\hat{\tau}_{s(k)}$.
%\item We rebuild
%\begin{eqnarray*}
%\hat{t}_{is(k)} &= &\hat{t}_{i}+\hat{t}_{s(k)} \\
%\hat{\tau}_{is(k)} &= & \hat{\tau}_{i}\times\hat{\tau}_{s(k)}
%\end{eqnarray*}
%\item We deduce the weighted average value of each component $\bar{t}$, $\bar{\tau}$ by year and transport mode, the weighting scheme being the relative value of the flows for each $i,s$ flow, as well as the median, the maximum and the minimum values.
%\item With this in hand, we can (among other things) obtain the estimated share of additive costs in total transport costs for each $i,k$ flow:
%
%
%$$\hat{\beta}_{ik} = \frac{\frac{\hat{t}_{is(k)}}{\tilde{p}_{ik}}}{\tau_{is(k)}-1 + \frac{t_{is(k)}}{\tilde{p}_{is(k)}}} $$
%\end{itemize}
%
%\paragraph{What the referee suggests to do}
%
%The estimation strategy suggested by the referee starts from Equation (\ref{eq:Hummels}) linking transport costs and unit prices as assumed in Hummels (2007). As shown in Subsection (\ref{ssec:interpret_beta}), the elasticity of transport costs to unit prices $\beta_{ik}$ (in absolute value) also corresponds to the share of additive costs in total transport costs. The share of additive costs can hence be uncovered by regressing transport costs on unit (ie, fas) prices. Specifically, the referee suggests to run the estimation based on Equation (\ref{eq:estimation_ref1}).
%
%Precisely, by year and transport mode, this implies running the estimation for each country of origin $i$ and each sector $s(k)$ = 3 digit sector, exploiting the variability between sub-sectors at the 10 digit level ($k$) and between ports of entry in the US ($d$).\footnote{The referee suggests to run his/her proposed estimation strategy considering sectoral aggregation at the 5-digit level (with $k=10$). We followed his suggestion but considering the 3-digit level. We made this choice in view of being able to compare the pros and cons of the referee's method versus ours. In Hummels' (2007) database, which we use until 2004, the finest degree of classification is at the 5-degree level until 2004. For comparability reasons over time, we adopt the same degree of classification $k=5$ for the following years (based on the original Census dataset). As a result, transport costs are estimated at the 3-digit level as benchmark estimate, exploiting heterogeneity at the 5-digit level (for a given origin country, and conditional on year and transport mode). This has driven us to retain $s=3$ when implementing the referee's method, in view of preserving comparability between both strategies.}

@@

\paragraph{a) The advantages of the referee's method are not as specific to this method as suggested}

\begin{itemize}
\item \textit{The first advantage of this so-called Hummels' Approach is that the above
regression can be estimated separately for various country-industry pairs, without
imposing Assumptions (a) $\tau_{ik} = \tau_i\tau_{k}$, and (b) $t_{ik} = t_i+ t_{k}$.} \\

\textbf{Our answer:}
As already discussed in this answer, the separability assumption does not seem to be a strong assumption. This is based on the conclusion drawn from the robustness check on a reduced sample. As we conclude at the end of Section 3.3.1 \textbf{XXX CHECK XXX}, whatever the transport mode and for both types of transport costs, the trend patterns of international transport costs are very similar whether estimated under the separability assumption or not.

@@

\item \textit{The second advantage is that there is a handful of previously-proposed instruments (e.g., HS-10 product-specific tariff rates or lagged prices), which the authors can use to overcome the endogeneity problem.}

     \textbf{Our answer:}
We thank the referee for this valuable suggestion. It is worth noticing that we can also handle the instrumentation of fas prices at the HS-10 level with our method - which we do in the revised version.

\item  \textit{The third advantage is that, by adopting this approach, the comparison between the paper [...] and those in Hummels (2007) would become more transparent.}\\
     \textbf{Our answer:}
     We agree with the referee that the comparison with the literature (\citealp{hummels2007} in particular) was not straightforward for the reader in the submitted version. We did our best to make the comparison clearer in the revised version, in a way consistent with our estimation method. In contrast to the submitted version, the paper now starts from the relation linking transport costs to the unit price specified in Hummels (2007):
    \begin{equation}
    f_{ikt} = X_{is(k)t}\widetilde{p}_{ikt}^{-\beta_{ikt}} \label{eq:Hummels}
    \end{equation}

     with $f_{ikt} = \frac{p_{ikt}}{\widetilde{p}_{ikt}} -1 $ the transport costs measure, $p_{ikt}$ ($\widetilde{p}_{ikt}$) the cif (fas) price and $\beta_{ikt}$ the price-elasticity of transport costs, with $i$ the origin country, $k$ the product and $t$ for time. As we now show in the paper (Section 2.2), $\beta_{ikt}$ also corresponds to the share of additive costs in total transport costs:\footnote{For sake of expositional purpose, the reasoning is made without any $i,k,t$ specification in the revised paper. We directly report them here to ease the comparison with the referee's method afterwards.}

@@

\begin{eqnarray}
p_{ik} &=& \tau_{is(k)}\widetilde{p}_{ik} +t_{is(k)} \label{eq:system1}\\
\frac{t_{is(k)}}{(\tau_{is(k)}-1)\widetilde{p}_{ik}+ t_{is(k)}} &=& \widehat{\beta}_{is(k)}  \label{eq:system2}
\end{eqnarray}

\noindent with $p_{ik}$ and $\widetilde{p}_{ik}$ the cif and fas prices observed in our dataset (conditional on a given year-transport mode). With two equations and two endogenous variables, the system can be solved.

 Alternatively, our method rather starts from the definition of transport costs as:
     $$\ln f_{ik} = \ln\left(\tau_{is(k)} -1 + \frac{t_{is(k)}}{\widetilde{p}_{ik}}\right) +\varepsilon_{ik}$$
\noindent from which we deduce the additive/multiplicative costs $\widehat{t}_{is(k)}$, $\widehat{\tau}_{is(k)}$ (on a mode/yearly basis). Then, relying on the Hummels's equation, one can deduce the share of additive costs in total transport costs $\beta_{ik}$ through Equation (\ref{eq:beta_TC}), again on a mode/yearly basis.

In this respect, the referee's method and ours are equivalent, in that they both allow to uncover the share of additive costs in total transport costs $\beta$, as well as the value of each trade costs component ($t,\tau$), that vary over time, country, sector and transport mode. Put it differently, our method is consistent with Hummels' methodology - provided that it is adequately reported as such. In comparison with \cite{hummels2007}, one supplementary advantage of our method is that we estimate a $\beta$ that varies over time, origin country and sector. Section 4 ([Section 3] of the submitted version) explores the importance of allowing for such a variability in accounting for the sources of the transport costs time trend. We agree that this was not sufficiently transparent in the submitted version. We hope that the revised version now fulfills the referee's expectations on this point.

@@

\end{itemize}

As previously noted, running this estimation limits the time coverage as information is available at the HS-10 level and by ports of entry only since 1989 (though we could not obtain the whole data because of  COVID). We have thus run referee's method over the years 2005-2013, at the $s=3, k=10$ sectoral/product levels, and reported the results in the Online Appendix, Section D.3, Table D.1 and Figure D.3.\textbf{XXX Check XXX}\footnote{Notice that we are aware of the referee's suggestion to run his/her estimation strategy considering sectoral aggregation at the 5-digit level (with $k=10$). We yet keep considering sectors at the 3-digit aggregation level, in view of being able to compare the referee's method versus ours.} For sake of reading clarity, we also report here the results of Table D.1 (here Table \ref{tab_comp_referee1_baseline}).

@@

\begin{table}[htbp]
	\caption{Comparison 2005-2013}
	\begin{center}		
		\input{../../revised_article/Online_Appendix/Comp_baseline_referee1_reshaped.tex}
		%\input{../../revised_article/Online_Appendix/Comp_baseline_referee1.tex}

{\parbox[l]{12cm}{ \vspace{4pt}\footnotesize{Notes: Time trend coefficient is the annual growth rate.}}}
\end{center}
	\label{tab_comp_referee1_baseline}%
\end{table}%

Regarding the share of additive costs $\beta$, the comparison between the two methods goes in opposite directions depending on the transport mode. While the estimated $\beta$ is lower for vessel transport with the alternative method (also displaying a larger decrease over the period), the opposite holds for Air transport. In any case, the share of additive costs remains substantial, around 40\% on average over the period. In this respect, it confirms the main message of our paper shedding light on the importance of the additive component in international transport costs.

@@

Yet, the suggested method does not go without a few important limitations, that we now develop.

\paragraph{b) The referee's method is not without drawbacks}

\begin{itemize}
\item[Concern 1] The suggested estimation strategy implies having much less data to cover.

### 2021-11-23 redaction/JEGeo/revision_JEGeo/revised_article/Online_Appendix/online_appendix_DHP_rev1.tex
petite coquille OA
Modification des tableaux de l’appendice
Signed-off-by: Guillaume Daudin <gdaudin@mac.com>
Trois des tableaux supplémentaires apparaissent
Signed-off-by: Guillaume Daudin <gdaudin@mac.com>
Les tableaux sont là
Signed-off-by: Guillaume Daudin <gdaudin@mac.com>
Suite au mail de Jérôme

This section is devoted to presenting the results. We first compare the estimates of the transport costs components under the three models. In a second step, we report quality of fit tests for the three models.

@@

In this section, we report the estimation results of the three models: Model (A) (with only ad-valorem costs), Model (B) (with both additive and ad-valorem) and Model (C) (with only additive costs). Table \ref{tab:3models_estimation_results_air} reports the results for air transport, Table \ref{tab:3models_estimation_results_ves} for vessel transport.

@@

Tables \ref{tab:result_air_rob} and \ref{tab:result_ves_rob} report the estimates of both models (with and without additive costs) in air and ocean transport respectively. Estimates at the 4-digit level confirm the significant role of the additive component in fitting international transport costs, already established at the 3-digit level.

@@

@@

@@

\begin{table}[htbp]
  \centering
  \caption{Vessel: Measures of goodness of fit, 4-digit level}
\begin{center}
\label{tab:good_fit_ves_rob}%
  \footnotesize{
\begin{tabular}{l|cccccc}
\hline
\hline
      & \multicolumn{6}{c}{Year}                   \\
      & \multicolumn{1}{c}{1974} & \multicolumn{1}{c}{1981} & \multicolumn{1}{c}{1989} & \multicolumn{1}{c}{2001} & \multicolumn{1}{c}{2009} & \multicolumn{1}{c}{2013}  \\ \hline
\multicolumn{7}{l}{\textbf{R$^{2}$}} \\ \hline
Model (A) & 0.50  & 0.45  & \multicolumn{1}{c}{0.47} & \multicolumn{1}{c}{0.41} & \multicolumn{1}{c}{0.37} & \multicolumn{1}{c}{0.35} \\
Model (B) \& I & 0.66  & 0.62  & \multicolumn{1}{c}{0.62} & \multicolumn{1}{c}{0.58} & \multicolumn{1}{c}{0.51} & \multicolumn{1}{c}{0.46} \\ \hline
\multicolumn{7}{l}{\textbf{SER} }\\ \hline
Model (A) &    0.58   &   0.64    & \multicolumn{1}{c}{0.61} & \multicolumn{1}{c}{0.0.72} & \multicolumn{1}{c}{0.79} & \multicolumn{1}{c}{0.82} \\
Model (B) &   0.48    &    0.53   & \multicolumn{1}{c}{0.51} & \multicolumn{1}{c}{0.61} & \multicolumn{1}{c}{0.69} & \multicolumn{1}{c}{0.75} \\ \hline
\multicolumn{7}{l}{\textbf{Log-likelihood} } \\ \hline
Model (A)& -16460.1 & -16951.6 & \multicolumn{1}{c}{-26771.4} & \multicolumn{1}{c}{-39008.3} & \multicolumn{1}{c}{-43888.9} & \multicolumn{1}{c}{-47161.6} \\
Model (B) & -12743.65 & -13546.9 & \multicolumn{1}{c}{-21752.8} & \multicolumn{1}{c}{-33281.0} & \multicolumn{1}{c}{-39078.9} & \multicolumn{1}{c}{-43399.2}  \\ \hline
\multicolumn{7}{l}{\textbf{AIC criteria} } \\ \hline
Model (A) & 34464.2 & 35491.2 & \multicolumn{1}{c}{55272.9} & \multicolumn{1}{c}{79800.7} & \multicolumn{1}{c}{89459.8} & \multicolumn{1}{c}{95987.2}\\
Model (B) & 28271.3 & 29877.8 & \multicolumn{1}{c}{46595.6} & \multicolumn{1}{c}{69743.9} & \multicolumn{1}{c}{81155.7} & \multicolumn{1}{c}{89692.4} \\ \hline
\multicolumn{7}{l}{\textbf{Test LL}} \\ \hline
Stat LL ratio (B vs A) & 12385.80 & 11226.8 & \multicolumn{1}{c}{17354.7} & \multicolumn{1}{c}{20113.5} & \multicolumn{1}{c}{16608.2} & \multicolumn{1}{c}{12589.6} \\
\# of restrictions (B vs A) & 797   & 814   & \multicolumn{1}{c}{881} & \multicolumn{1}{c}{910} & \multicolumn{1}{c}{886} & \multicolumn{1}{c}{874} \\
p-value (B vs A) & 0.000 & 0.000 & \multicolumn{1}{c}{0.000} & \multicolumn{1}{c}{0.000} & \multicolumn{1}{c}{0.000} & \multicolumn{1}{c}{0.000} \\
\hline\hline
\multicolumn{7}{l}{\parbox[l]{13cm}{ \vspace{7pt}\scriptsize{Notes: Model (A) = with only ad-valorem transport costs.
Model (B) = with additive \& ad-valorem transport costs.
$R^{2}$ between the log of predicted ratio and the log of the observed ratio.
The number \# of restrictions is equal to the number of parameters estimated, i.e. the number of partner countries plus the number of products.}}}
\end{tabular}%
}
\end{center}
\end{table}%

If anything, the quality of fit appears slightly higher when estimations are based on the 4-digit classification. This is especially true for the model restricting transport cost to their ad-valorem dimension, whatever the transport mode considered.
When the additive part is taken into account however, the difference in goodness of fit between the 3- and the 4-digit classification level becomes very small, whatever the considered criterion.
In other words, if using a more disaggregated product classification unsurprisingly adds some statistical precision, this is not to an extent that would disqualify the use of slightly more aggregated data.

@@

\subsection{Transport Cost Estimates: Yearly Detailed Results}

@@

		\input{TableB1974_1987_ves.tex}
				%\begin{tablenotes}
%			\tiny
%			\item
%			\item \textbf{Model (A): With ad-valorem transport costs only}
%			\item \textbf{Model (B): With additive and ad-valorem transport costs}
%			\item \textbf{Model (C): With additive transport costs only}
%		\end{tablenotes}

@@

		\input{TableB1988_2001_ves.tex}
				%\begin{tablenotes}
%			\tiny
%			\item Statistics are weighted by value
%			\item \textbf{Model (A): With ad-valorem transport costs only}
%			\item \textbf{Model (B): With additive and ad-valorem transport costs}
%			\item \textbf{Model (C): With additive transport costs only}
%		\end{tablenotes}

@@

		\input{TableB2002_2015_ves.tex}
				%\begin{tablenotes}
%			\tiny
%			\item Statistics are weighted by value
%			\item \textbf{Model (A): Iceberg transport costs only}
%			\item \textbf{Model (B): With additive and ad-valorem transport costs}
%			\item \textbf{Model (C): With additive transport costs only}
%		\end{tablenotes}

@@

		\input{TableB2016_2019_ves.tex}
			%	\begin{tablenotes}
%			\tiny
%			\item Statistics are weighted by value
%			\item \textbf{Model (A): Iceberg transport costs only}
%			\item \textbf{Model (B): With additive and ad-valorem transport costs}
%			\item \textbf{Model (C): With additive transport costs only}
%		\end{tablenotes}

@@

As mentioned in the paper, the estimates for air transport costs in 1989 show a surprisingly high value for the additive component (the additive cost is estimated to amount to $4.6\%$ of the export price, whereas it amounts to 2.5\% on average between 1974 and 1988, and to 1.7\% over the following decade 1990-2000). This can be attributed to the presence of outliers in the distribution of the additive costs estimates. The maximum value for $\widehat{t}/\widetilde{p}$ is 10,000\% in 1989, whereas it amounts to 1,690\% on average over 1974-1988 and to 1,500\% on average over 1990-2000. Accordingly, in the paper we discard this year 1989 when we report the average values over the period of the transport costs estimates in air transport.\\
Still, this does not make much difference. When 1989 is included, the weighted mean transport cost value amounts to 1.9\% for the ad-valorem component, vs 1.8\% when 1989 is excluded. The weighted median value is left unchanged at 2.9\% whether 1989 is included or not.

@@

		\input{TableB1974_1987_air.tex}
				%\begin{tablenotes}
%			\tiny
%			\item Statistics are weighted by value
%			\item \textbf{Model (A): With ad-valorem transport costs only}
%			\item \textbf{Model (B): With additive and ad-valorem transport costs}
%			\item \textbf{Model (C): With additive transport costs only}
%		\end{tablenotes}
%		
	\end{center}
	\label{tab_oa:result_air_ally3}%
\end{table}%

>>>>

In this section, we first compare the estimates of the transport costs components under the three models. In a second step, we report quality of fit tests for the three models.

@@

We report estimations of the three models for selected years. Section \ref{secoa:yearly_results} provides estimations for every year. Table \ref{tab:3models_estimation_results_air} reports the results for air transport, Table \ref{tab:3models_estimation_results_ves} for vessel transport.

@@

Tables \ref{tab:result_air_rob} and \ref{tab:result_ves_rob} report the estimates of both models (with and without additive costs) in air and vessel transport respectively. Estimates at the 4-digit level confirm the significant role of the additive component in fitting international transport costs we found at the 3-digit level.

@@

\begin{table}[htbp]
	\centering
%	\footnotesize{
		\caption{Air: Transport costs estimates, selected years, 4-digit level}
		\label{tab:result_air_rob}%
		\input{TableB1_air.tex}
		\begin{tablenotes}
			\tiny
			\item Statistics are weighted by value
			\item \textbf{Model (A): Iceberg transport costs only}
			\item \textbf{Model (B): With additive and ad-valorem transport costs}
		\end{tablenotes}
\end{table}%

@@

\begin{table}[htbp]
	\centering
%	\footnotesize{
		\caption{Vessel: Transport costs estimates, selected years, 4-digit level}
		\label{tab:result_air_rob}%
		\input{TableB1_ves.tex}
		\begin{tablenotes}
			\tiny
			\item Statistics are weighted by value
			\item \textbf{Model (A): Iceberg transport costs only}
			\item \textbf{Model (B): With additive and ad-valorem transport costs}
		\end{tablenotes}
\end{table}

@@

\begin{table}[htbp]
	\centering
%	\footnotesize{
		\caption{Air: Measures of goodness of fit, 4-digit level}
		\label{tab:good_fit_air_rob}%
		\input{TableB3_air.tex}
		\begin{tablenotes}
			\tiny
			\item Statistics are weighted by value
			\item \textbf{Model (A): Iceberg transport costs only}
			\item \textbf{Model (B): With additive and ad-valorem transport costs}
		\end{tablenotes}
\end{table}

\begin{table}[htbp]
  \centering
  \caption{Vessel: Measures of goodness of fit, 4-digit level}
\begin{center}
\label{tab:good_fit_ves_rob}%
  \footnotesize{
\begin{tabular}{l|cccccc}
\hline
\hline
      & \multicolumn{6}{c}{Year}                   \\
      & \multicolumn{1}{c}{1974} & \multicolumn{1}{c}{1981} & \multicolumn{1}{c}{1989} & \multicolumn{1}{c}{2001} & \multicolumn{1}{c}{2009} & \multicolumn{1}{c}{2013}  \\ \hline
\multicolumn{7}{l}{\textbf{R$^{2}$}} \\ \hline
Model (A) & 0.50  & 0.45  & \multicolumn{1}{c}{0.47} & \multicolumn{1}{c}{0.41} & \multicolumn{1}{c}{0.37} & \multicolumn{1}{c}{0.35} \\
Model (B) & 0.66  & 0.62  & \multicolumn{1}{c}{0.62} & \multicolumn{1}{c}{0.58} & \multicolumn{1}{c}{0.51} & \multicolumn{1}{c}{0.46} \\ \hline
\multicolumn{7}{l}{\textbf{SER} }\\ \hline
Model (A) &    0.58   &   0.64    & \multicolumn{1}{c}{0.61} & \multicolumn{1}{c}{0.0.72} & \multicolumn{1}{c}{0.79} & \multicolumn{1}{c}{0.82} \\
Model (B) &   0.48    &    0.53   & \multicolumn{1}{c}{0.51} & \multicolumn{1}{c}{0.61} & \multicolumn{1}{c}{0.69} & \multicolumn{1}{c}{0.75} \\ \hline
\multicolumn{7}{l}{\textbf{Log-likelihood} } \\ \hline
Model (A)& -16460.1 & -16951.6 & \multicolumn{1}{c}{-26771.4} & \multicolumn{1}{c}{-39008.3} & \multicolumn{1}{c}{-43888.9} & \multicolumn{1}{c}{-47161.6} \\
Model (B) & -12743.65 & -13546.9 & \multicolumn{1}{c}{-21752.8} & \multicolumn{1}{c}{-33281.0} & \multicolumn{1}{c}{-39078.9} & \multicolumn{1}{c}{-43399.2}  \\ \hline
\multicolumn{7}{l}{\textbf{AIC criteria} } \\ \hline
Model (A) & 34464.2 & 35491.2 & \multicolumn{1}{c}{55272.9} & \multicolumn{1}{c}{79800.7} & \multicolumn{1}{c}{89459.8} & \multicolumn{1}{c}{95987.2}\\
Model (B) & 28271.3 & 29877.8 & \multicolumn{1}{c}{46595.6} & \multicolumn{1}{c}{69743.9} & \multicolumn{1}{c}{81155.7} & \multicolumn{1}{c}{89692.4} \\ \hline
\multicolumn{7}{l}{\textbf{Test LL}} \\ \hline
Stat LL ratio (B vs A) & 12385.80 & 11226.8 & \multicolumn{1}{c}{17354.7} & \multicolumn{1}{c}{20113.5} & \multicolumn{1}{c}{16608.2} & \multicolumn{1}{c}{12589.6} \\
\# of restrictions (B vs A) & 797   & 814   & \multicolumn{1}{c}{881} & \multicolumn{1}{c}{910} & \multicolumn{1}{c}{886} & \multicolumn{1}{c}{874} \\
p-value (B vs A) & 0.000 & 0.000 & \multicolumn{1}{c}{0.000} & \multicolumn{1}{c}{0.000} & \multicolumn{1}{c}{0.000} & \multicolumn{1}{c}{0.000} \\
\hline\hline
\multicolumn{7}{l}{\parbox[l]{13cm}{ \vspace{7pt}\scriptsize{Notes: Model (A) = with only ad-valorem transport costs.
Model (B) = with additive \& ad-valorem transport costs.
$R^{2}$ between the log of predicted ratio and the log of the observed ratio.
The number \# of restrictions is equal to the number of parameters estimated, i.e. the number of partner countries plus the number of products.}}}
\end{tabular}%
}
\end{center}
\end{table}%

\begin{table}[htbp]
	\centering
%	\footnotesize{
		\caption{Vessel: Measures of goodness of fit, 4-digit level}
		\label{tab:good_fit_air_rob}%
		\input{TableB3_ves.tex}
		\begin{tablenotes}
			\tiny
			\item Statistics are weighted by value
			\item \textbf{Model (A): Iceberg transport costs only}
			\item \textbf{Model (B): With additive and ad-valorem transport costs}
		\end{tablenotes}
	\end{table}

If anything, the quality of fit appears slightly higher when estimations are based on the 4-digit classification. This is especially true for the model restricting transport cost to their ad-valorem dimension, whatever the transport mode considered.
When the additive part is taken into account however, the difference in goodness of fit between the 3- and the 4-digit classification level becomes very small, whatever the considered criterion.
Using a more disaggregated product classification unsurprisingly adds some statistical precision, but the gain is not large enough to disqualify the use of slightly more aggregated data.

@@

\subsection{Transport Cost Estimates: Yearly Detailed Results \label{secoa:yearly_results}}

@@

		\input{TableC1974_1987_ves.tex}
				%\begin{tablenotes}
%			\tiny
%			\item
%			\item \textbf{Model (A): With ad-valorem transport costs only}
%			\item \textbf{Model (B): With additive and ad-valorem transport costs}
%			\item \textbf{Model (C): With additive transport costs only}
%		\end{tablenotes}

@@

		\input{TableC1988_2001_ves.tex}
				%\begin{tablenotes}
%			\tiny
%			\item Statistics are weighted by value
%			\item \textbf{Model (A): With ad-valorem transport costs only}
%			\item \textbf{Model (B): With additive and ad-valorem transport costs}
%			\item \textbf{Model (C): With additive transport costs only}
%		\end{tablenotes}

@@

		\input{TableC2002_2015_ves.tex}
				%\begin{tablenotes}
%			\tiny
%			\item Statistics are weighted by value
%			\item \textbf{Model (A): Iceberg transport costs only}
%			\item \textbf{Model (B): With additive and ad-valorem transport costs}
%			\item \textbf{Model (C): With additive transport costs only}
%		\end{tablenotes}

@@

		\input{TableC2016_2019_ves.tex}
			%	\begin{tablenotes}
%			\tiny
%			\item Statistics are weighted by value
%			\item \textbf{Model (A): Iceberg transport costs only}
%			\item \textbf{Model (B): With additive and ad-valorem transport costs}
%			\item \textbf{Model (C): With additive transport costs only}
%		\end{tablenotes}

@@

As mentioned in the paper, the estimates for air transport costs in 1989 show a surprisingly high value for the additive component (the additive cost is estimated to amount to $4.6\%$ of the export price, whereas it amounts to 2.5\% on average between 1974 and 1988, and to 1.7\% over the following decade 1990-2000). This can be attributed to the presence of outliers in the distribution of the additive costs estimates, despite the exclusion of outliers in the data. The maximum value for $\widehat{t}/\widetilde{p}$ is 10,000\% in 1989, whereas it amounts to 1,690\% on average over 1974-1988 and to 1,500\% on average over 1990-2000. Accordingly, in the paper we discard this year 1989 when we report the average values over the period of the transport costs estimates in air transport.\\
Excluding or excluding a single year out of 45 does not make much difference in the weighted mean transport cost.
%When 1989 is included, the weighted mean transport cost value amounts to 1.9\% for the ad-valorem component, vs 1.8\% when 1989 is excluded. The weighted median value is left unchanged at 2.9\% whether 1989 is included or not.

@@

		\input{TableC1974_1987_air.tex}
				%\begin{tablenotes}
%			\tiny
%			\item Statistics are weighted by value
%			\item \textbf{Model (A): With ad-valorem transport costs only}
%			\item \textbf{Model (B): With additive and ad-valorem transport costs}
%			\item \textbf{Model (C): With additive transport costs only}
%		\end{tablenotes}
%		
	\end{center}
	\label{tab_oa:result_air_ally3}%
\end{table}%
